---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {rmlx.design} - Structured data splitting to facilitate estimand-aligned performance estimation in machine learning (R package)





The {rmlx.design} package allows structured data splitting for supervised machine learning tasks. 
The data splitting in {rmlx.design} is based on an **estimand** definition which is in turn based
on user-specified constraints.

**Constraints** can be thought of as generalized inclusion/exclusion criteria for 

1. Test observations
2. The relation between observations in training and test sets
3. The training dataset(s)

In contrast to traditional techniques (hold-out, cross-validation, bootstrap), this leads to a
deterministic data splitting. The intention behind structured data splitting is to allow estimation of **estimand-aligned transferabiliy (out-of-distribution generalization)** instead of **reproducibility (in distribution generalization)** (Alpers & Westphal, 2025).


## Installation

You can install the development version of {rmlx.design} from [GitHub](https://github.com/) with:

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("maxwestphal/rmlx.design")
```

## Getting started

An overview over important functions and classes and a minimal example can be found in the "overview" vignette:

```{r vignette, eval=FALSE}
library(rmlx.design)
vignette("overview", "rmlx.design")
```


## Roadmap

Planned features for future versions include:

1. Extension of traditional (unstructured) data splitting methods 
2. Export of data splitting schemes to popular ML packages
3. Improved summary/visualization functions for derived data splits


## References

- Alpers, R. and Westphal, M. 2025. An estimand framework to guide model and
algorithm validation in predictive modelling. Submitted for publication.
