---
title: "{rmlx.design} - package overview"
output: rmarkdown::html_vignette
always_allow_html: true
vignette: >
  %\VignetteIndexEntry{overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
library(kableExtra)
library(rmlx.design)
library(mlr3verse)
library(ggplot2)
```


```{r, include=FALSE}
options(kableExtra.html.bsTable = TRUE)

tab <- function(x) {
  x %>%
    kbl() %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
    scroll_box(height = "400px")
}

tab_splits <- function(x) {
  x$info %>% tab()
}
```



## Main functionality

The {rmlx.design} package allows meaningful data splitting for supervised machine learning tasks. 
The data splitting in {rmlx.design} is based on an **estimand** definition which is in turn based
on user-specified constraints.

**Constraints** can be thought of generalized inclusion/exclusion criteria for 

1. Test observations
2. The relation between observations in training and test sets
3. The training dataset(s)

In contrast to traditional techniques (hold-out, cross-validation, bootstrap), this leads to a
deterministic data splitting. The following diagram shows important relationships between objects (of relevant classes) and functions implemented in the {rmlx.design} package.

```{r, setup2, include=FALSE}
library(nomnoml)
```

```{nomnoml, echo=FALSE, svg=TRUE, width="8in", height="8in"}
#direction: down
#stroke: black
#.frame: bold
#.obj: stroke=blue
#.fun: stroke=#ff8c00 visual=roundrect bold


[<frame> mldesign workflow for estimand-based data splitting|
  [<fun> specify_constraint : function]
  [<fun> specify_constraint : function] out -> [c1: rmlx_constraint]
  [<fun> specify_constraint : function] out -> [c2: rmlx_constraint]
  [<fun> specify_constraint : function] out -> [c3: rmlx_constraint]
  [c1: rmlx_constraint | name = "ic_age" | expr = "test$age >= 18" | vars = "age" | type = "population" | target = "test"]
  [c2: rmlx_constraint | name = "temporal_transferability" | expr = "test$year - train$year >= 1" | vars = "year" | type = "context" | target = "relation"]
  [c3: rmlx_constraint | name = "sufficient n_train" | expr = "nrow(train) >= 100" | vars = "" | type = "admissibility" | target = "train"]
  [c1: rmlx_constraint] in --> [<fun> specify_estimand : function]
  [c2: rmlx_constraint] in --> [specify_estimand : function]
  [c3: rmlx_constraint] in --> [specify_estimand : function] 
 
  [specify_estimand : function] out -> [my_estimand: rmlx_estimand | test = list(c1, ...) |  relation = list(c2, ...) |  train = list(c3, ...) ]
 
  [my_data: data.frame] in --> [<fun> derive_splits : function]
  [my_estimand: rmlx_estimand] in --> [my_design: rmlx_design]
  [my_design: rmlx_design] --> [derive_splits : function]
 
  [derive_splits : function] out -> [my_splits: rmlx_splits]
 
  [my_data: data.frame] in --> [<fun> get_data_split : function]
  [my_splits: rmlx_splits | list(\n list(train, test),\n list(train, test),\n ...\n) ] in --> [<fun> get_data_split : function]
  [index: integer] in --> [<fun> get_data_split : function]
  
  [get_data_split : function] out -> [my_data_split: list | train: data.frame | test: data.frame]
 
]
```

## Custom classes

1. **rmlx_constraint**: a single constraint
1. **rmlx_constraints**: a collection of constraints
1. **rmlx_estimand** (subclass of rmlx_design): three constraints objects *test*, *relation*, *train*
1. **rmlx_splitting** (subclass of rmlx_design): defines a traditional data splitting method (hold-out, cross-validation, ...)
1. **rmlx_splits**: a list of all train/test indices


## Important functions

1. **parse_constraint()**: to define a constraint only based on a single string
1. **specify_constraint() / sc()**: to define a constraint
1. **combine_constraints() / cc()**: to combine multiple constraints
1. **specify_estimand()**: to define an estimand
1. **specify_splitting_rule()**: to define a data splitting method
1. **nest()**: to define a nested data splitting scheme
1. **derive_splits(spec, data)**: to conduct the data splitting based on the estimand specification, returns indices only
1. **split_data(splits, data, idx)**: to get a single train/set split, return data sets (subsets of original data)







## Example

### Constraints

Generate some data for demonstration purposes
```{r}
set.seed(1337)
data <- generate_data(n_obs = 2000)
names(data)
head(data)
```

Imagine we would want to apply our (trained) prediction model to predict the outcome variable only 
to patients older than 50 years:

```{r}
specify_constraint(~ test$age > 50)
```
Note that variables in this expression will automatically be recognized as long as the syntax 
`test[['age']]` (preferred) or `test$age` (alternative, can cause problems in some cases) is
used. Currently, detection of variables is still in an early stage via `detect_vars()` 
but can be overwritten via (e.g.) `specify_constraint(..., vars = "age")`.

We can define further (optional) information. This will not affect results down the line but 
will improve documentation
```{r}
specify_constraint(~ test$age > 50, name = "ic_age", type = "population")
```

Note that we can specify the same constraint in a single character object with help of the parse_constraint function:

```{r}
"test[['age']] > 50 #name=ic_age #type=population" %>% parse_constraints()
```
Note that the use of the "#" character to specify further argument of specify_constraint().

The function combine_constraints() (shortcut: cc()) can be used to bundle multiple rmlx_constraint
objects into a single rmlx_constraints object. 
It can also handle character inputs (via parse_constraint) or rmlx_constraints inputs.


### Estimand specification

To define a first, simple estimand, let's focus on the *relation* constraints. These constraints 
define the relationship between test and training observations. In this case, we specify constraints
that we expect to hold between training dataset **train** and testing dataset **test**. Note, that these
datasets do not yet exist explicitly but rather will be derived later 
(possibly in multiple versions) based on our estimand
specification and our entire dataset **data**.


```{r}
constraints_relation <- cc(
  specify_constraint(~ (test[["year"]] - train[["year"]]) %in% 1:2,
    name = "time_transferability", type = "context"
  ),
  specify_constraint(~ test[["country"]] != train[["country"]],
    name = "region_transferability", type = "context"
  )
)
```

```{r}
constraints_relation %>% str(2)
```
```{r}
my_estimand <- specify_estimand(
  constraints = constraints_relation,
  name = "my_estimand"
)
my_estimand
```

```{r}
my_design <- specify_design(my_estimand)
my_design
```



### Data splitting 

Now, we use the estimand to conduct (deterministically) the corresponding data splitting:

```{r}
splits <- derive_splits(my_design, data)
```

```{r}
tab_splits(splits)
```
As we can see, we obtain a list of train/test indices corresponding to the estimand specification.
To finally obtain concrete train/test sets (the first one out of the overall list):

```{r}
split_data(data, splits, idx = 1) %>% str()
```



### Nested data splitting

It is also possible to define nested data splitting schemes that are (partially) structured via the
**nest()** function or its alias **specify_nested()**. The result will be similar to the well-known 
nested cross-validation technique. In fact, let's start by defining the conventional (5 x 5) nested CV to start


```{r}
nest(
  specify_splitting_rule("cv", n_folds = 5)
) %>%
  specify_design() %>% 
  derive_splits(data) %>%
  tab_splits()
```

By default, the outer data splitting is also used for the inner data splitting, if not specified otherwise.

In addition, we can define a structured nested scheme as follows:

```{r}
nest(
  my_estimand,
  specify_splitting_rule("cv", n_folds = 3)
) %>%
  specify_design() %>% 
  derive_splits(data) %>%
  tab_splits()
```

So this splitting scheme is structured in the outer loop to assess the estimand-aligned transferability and
based on standard (3-fold) cross-validation in the inner loop for assessment of reproducibility 
(and usually also for hyperparameter tuning).




### Export to {mlr3}

After defining the relevant estimand(s), of course we want to do some actual machine learning.
Let's do this within the great [mlr3](https://mlr3.mlr-org.com/) framework. To achieve this, 
we will utilize the custom resampling scheme in {mlr3} via `rsmp("custom")`.

```{r}
splits <-
  nest(
    specify_estimand(sc(~ test[["country"]] != train[["country"]])),
    specify_splitting_rule("cv", n_folds = 3)
  ) %>%
  specify_design() %>% 
  derive_splits(data) %>%
  add_mlr3_vars()

str(splits, 1)
```
The last computation `add_mlr3_vars` is optional but will save us some time later on
(otherwise, this computation will be done at least twice). In this step, the actual relevant train and
test sets for {mlr3} are derived. This will save some time compared just iterating over all mldesign splits,
because for some splits the training data is the same. In effect, we avoid training the same model
multiple times, at least for nested data splitting schemes. Let's check this:

```{r}
nrow(splits$info)
length(splits$mlr3$map)
```
This implies that there are `r nrow(splits$info)` splits here which are 'condensed' to `r `length(splits$mlr3$map)` training (and testing) in runs for mlr3.



```{r, message=FALSE}
task <- as_task_classif(data %>% dplyr::select(age, comorbidity, outcome),
  target = "outcome",
  positive = "1",
  id = "example_task"
)

exp_design <- rsmp("custom")
exp_design$instantiate(task,
  train_sets = get_mlr3_sets_train(splits),
  test_sets = get_mlr3_sets_test(splits)
)

tune_space <- mlr3tuningspaces::lts("classif.ranger.default")
learner <- tune_space$get_learner()
learner$predict_type <- "prob"
search_space <- learner$param_set$search_space()


## training:
set.seed(123)
suppressMessages({
  learner_tuned <- mlr3tuning::tune(
    tuner = tnr("random_search"),
    task = task,
    learner = learner,
    resampling = exp_design,
    measures = msrs("classif.acc"),
    term_evals = 5
  )
})
learner_tuned
```

Let's use the convenience function `get_mlr3_predictions` to retreive predictions and label for all observations in all splits in a single data.frame.

```{r}
predictions <- get_mlr3_predictions(splits = splits, instance = learner_tuned)
dim(predictions)
head(predictions)
```




From here, we can calculate any relevant metric of interest, e.g. the area under the ROC curve (AUC). 
In the next step, we then select models based on the aggregate performance in the inner loop of the
nested resampling.

```{r}
data_benchmark <-
  derive_metrics(
    predictions = predictions,
    splits = splits,
    auc = pROC::auc(truth, prob.1, direction = "<", levels = c("0", "1")) %>% as.numeric()
  ) %>%
  select_models(metric = "auc", aggregate = mean, select = max)

tab(data_benchmark)
```


To conclude, we make a simple visualization for the resulting benchmark data. 

```{r}
data_benchmark %>%
  ggplot(aes(type, auc, size = n_train, color = factor(idx_outer))) +
  geom_point() +
  labs(size = bquote(n[train]), color = bquote(idx[outer]))
```



